---
title: "4_accessible_data_representation"
format: html
editor: visual
---

Goals

## 4.1 Improve Data Visualizations


```{r}
#Imports

library(hellodatascience)
library(tidyverse)

```

```{r}
base_bar_plot <-
  ggplot(
    data = atus_college, 
    aes(
      x = employment, 
      fill = enrollment
    )
  ) +
  geom_bar(position = "dodge")

base_bar_plot
```

## 4.2 Labels

```{r}
labeled_bar_plot <- 
  base_bar_plot + 
  labs(
    title = "Student Count by Employment and Enrollment Status",
    subtitle = "Most students work part-time while enrolled full-time",
    x = "Employment Status",
    y = "Number of Students",
    fill = "Course Enrollment",
    caption = "Data source: American Time Use Survey 2024 (n = 312 students)"
  )

labeled_bar_plot
```

## 4.1.2 Themes

```{r}
labeled_bar_plot

improved_bar_plot <- 
  labeled_bar_plot +
  theme_minimal()

improved_bar_plot
```

Other themes include: 
* theme_bw()
* theme_light() 
* theme_dark()
* theme_minimal()
* theme_classic()

```{r}
improved_bar_plot <-
  improved_bar_plot + 
#manually adjust text size
  theme(
    plot.title = element_text(size = 18),
    plot.subtitle = element_text(size = 14),
    axis.title = element_text(size = 12), 
    axis.text = element_text(size = 10), 
    legend.title = element_text(size = 11),
    legend.text = element_text(size = 9), 
    plot.caption = element_text(size = 8),
    #Adjust Legend
    legend.position = "top"
  ) +
labs(title = "Student Count by Employment and\n Enrollment Status")

improved_bar_plot

```

## 4.1.3 Choice of Color

colors() lists all options R accepts
* but can also use HEX

Remember to pick colors that work for people who might be color blind

Okabe-Ito Color Pallete

```{r}
palette.colors(palette = "Okabe-Ito")
```


```{r}
improved_bar_plot <-
  improved_bar_plot +
    scale_fill_manual(
      values = c(
       "Full Time" = "#009E73", #R Studio previews the color for you
       "Part Time" = "#CC79A7" #Which is very cool
    )
  )

improved_bar_plot
```


## Additional Style Elements 

```{r}
simple_scatterplot <- 
  ggplot(
    data = atus_college,
    aes(
      x = time_alone,
      y = weekly_earnings
    )
  ) +
  geom_point() +
  labs(
    x = "Time Alone (minutes)",
    y = "Weekly Earnings ($)"
  ) +
  theme_bw()

simple_scatterplot
```

```{r}
colored_scatterplot <- 
  simple_scatterplot +
  geom_point(
    color = "aquamarine4", #Also works for color Name
    size = 4,
    shape = "square"
  )

colored_scatterplot
```

```{r}
ggplot(
  data = atus_college, 
  aes(x = employment)
) +
  geom_bar(fill ="purple")
```

## 4.1.5 Facets

```{r}
simple_scatterplot +
  facet_grid(rows = vars(employment), 
             cols = vars(enrollment))
```

##  4.2 Data Verbalization


using labels' "alt" creaes alt text 

```{r}
improved_bar_plot +
  labs(
    alt = "Barplot showing student count by employment and enrollment status. The y-axis shows number of students from 0 to 80, and the x-axis shows employment status (Full Time, Part Time, NA). Each employment status is broken down by full time or part time enrollment status. The plot displays six bars: students enrolled full time with full time employment (approximately 70 students), part time employment (apprximately 75 students), and unknown employment (approximately 80 students) as well as students enrolled part time with full time employment (approxmately 60 students), part time employment (approximately 18 students), and unknown employment (approximately 15 students). Data source: American Time Use Survey 2024 (n = 312 students)."
  )
```

##  4.2 Data Sonification

```{r}
x <- 1:10
y <- 1:10

simple_plot <-
  ggplot(
    data = data.frame(x = x, y = y),
    aes(x = x,  y = y)) +
  geom_point()

simple_plot
```

install.packages("package_name")
```{r}
#Allows you to "listen" to the data
sonify::sonify(x, y)
```


## Bonus - Can insert code in other languages like Python or SQL

```{python}
print("hello world")
```



The `echo: false` option disables the printing of code (only output is displayed).
